---
title: "16S_Nvec_10272019"
author: "Anthony_Bonacolta"
date: "10/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE, echo=FALSE}
library("phyloseq"); packageVersion("phyloseq")
library("ggplot2"); packageVersion("ggplot2")
library("ape"); packageVersion("ape")
library("vegan"); packageVersion("vegan")
library("dplyr")
library("scales")
library("grid")
library("reshape2")
library("edgeR")
library("plyr")
library("tidyr")
library("viridis")
library("DESeq2")
library("gridExtra")
```
# Load Data into Phyloseq
```{r,echo=FALSE}
SV <- read.table("/Users/anthonybonacolta/Box/Nvec_Microbiome/R_data/feature-tableNvec_new.txt", row.names = 1, header = TRUE)
tax <-as.matrix(read.table("/Users/anthonybonacolta/Box/Nvec_Microbiome/R_data/tax-no-MitoChNvec_ps_new.txt", row.names = 1, header = FALSE, sep = "\t"))
tree_file <-read_tree("/Users/anthonybonacolta/Box/Nvec_Microbiome/R_data/tree.nwk")
colnames(tax) <- c ("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")
tax[tax==""] <- NA
map <- read.table("/Users/anthonybonacolta/Box/Nvec_Microbiome/R_data/16S_metadata_tab_phyloseq_new_revised.txt", sep ="\t", row.names = 1, header = TRUE)
ps = phyloseq(otu_table(SV, taxa_are_rows=TRUE), 
               sample_data(map), 
               tax_table(tax),
                 phy_tree(tree_file))
ps
```

# Decontaminate
```{r}
library(decontam); packageVersion("decontam")
head(sample_data(ps))
sample_data(ps)$is.neg <- sample_data(ps)$Sample_or_Control == "Control Sample"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg")
table(contamdf.prev$contaminant)
```
**5 OTUs are negatives**
```{r}
head(which(contamdf.prev$contaminant))
```
identify as contaminants all sequences thare are more prevalent in negative controls than in positive samples
54- Actinomyces
75- Lactobacillus
95- Alcaligenes
265- Prevotellaceae
565- Porphyromonas

```{r}
contamdf.prev05 <- isContaminant(ps, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev05$contaminant)
```
### Remove contaminants

```{r}
ps.noncontam <- prune_taxa(!contamdf.prev05$contaminant, ps)
ps.noncontam
```
## Just make it the real one
```{r}
ps <- ps.noncontam
ps = subset_samples(ps, Sample_or_Control != "Control Sample")
ps
```


# PCoA Plots
```{r, echo=FALSE, warning=FALSE, message = FALSE}
ord.PCoA.brayP <- ordinate(ps, method="PCoA", distance="bray")
plot_ordination(ps, ord.PCoA.brayP, color="Tissue", label = "Sample")
PCoATissueWneg <- plot_ordination(ps, ord.PCoA.brayP, color="Tissue", label = "Sample")
svg(filename="/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/PCoATissue.svg")
plot(PCoATissueWneg)
dev.off()
```
# Species Richness
```{r, echo=FALSE, warning=FALSE, message = FALSE}
plot_richness(ps, x="Tissue", color="Tissue",
              measures=c("Simpson", "Shannon")) + geom_point()
richness <- plot_richness(ps, x="Tissue", color="Tissue",
              measures=c("Simpson", "Shannon")) + geom_point()
svg(filename="/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/richness.svg")
plot(richness)
dev.off()
jpeg(filename="/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/richness.jpeg")
plot(richness)
dev.off()
```

# Abundance Figures
## Phylum Level
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Nvec_phylum <- ps %>%
  tax_glom(taxrank = "Phylum") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.02) %>%                         # Filter out low abundance taxa
  arrange(Phylum)  
phylum_colors <- c(
  "#CBD588", "#5F7FC7", "orange","#DA5724", "#508578", "#CD9BCD",
   "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861"
)

Tissue.labs <- c("Capitulum", "Mesenteries", "Physa")
names(Tissue.labs) <- c("capitulum", "mesenteries", "physa")


ggplot(Nvec_phylum, aes(x = Individual, y = Abundance, fill = Phylum)) + 
  facet_grid(Tissue~., labeller = labeller(Tissue = Tissue.labs)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = phylum_colors) +
  scale_x_discrete(
    breaks = c("1", "2", "3", "4", "5", "6"),
    labels = c("1", "2", "3", "4", "5", "6"), 
    drop = FALSE
  ) +
  # Remove x axis title
  xlab("Individual") +
  #
  guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) +
  ylab("Relative Abundance (Phyla > 2%) \n") +
  ggtitle("Phylum Composition of Nematostella vectensis \n Bacterial Communities by Tissue Compartment")

```
```{r, echo=FALSE, include=FALSE, warning=FALSE, message = FALSE}
Phylum_comp <- ggplot(Nvec_phylum, aes(x = Individual, y = Abundance, fill = Phylum)) + 
  facet_grid(Tissue~., labeller = labeller(Tissue = Tissue.labs)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = phylum_colors) +
  scale_x_discrete(
    breaks = c("1", "2", "3", "4", "5", "6"),
    labels = c("1", "2", "3", "4", "5", "6"), 
    drop = FALSE
  ) +
  # Remove x axis title
  xlab("Individual") +
  #
  guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) +
  ylab("Relative Abundance (Phyla > 2%) \n") +
  ggtitle("Phylum Composition of Nematostella vectensis \n Bacterial Communities by Tissue Compartment")

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/Revised_Phylum_comp.tiff", units="in", width=5, height=5, res=300)
plot(Phylum_comp)
dev.off()
```
## Class
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Nvec_class <- ps%>%
  tax_glom(taxrank = "Class") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.02) %>%                         # Filter out low abundance taxa
  arrange(Class) 

class_colors <- c(
  "#CBD588", "#5F7FC7", "orange","#DA5724", "#508578", "#CD9BCD",
   "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861"
)
ggplot(Nvec_class, aes(x = Individual, y = Abundance, fill = Class)) + 
  facet_grid(Tissue~., labeller = labeller(Tissue = Tissue.labs)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = class_colors) +
  scale_x_discrete(
    breaks = c("1", "2", "3", "4", "5", "6"),
    labels = c("1", "2", "3", "4", "5", "6"), 
    drop = FALSE
  ) +
  # Remove x axis title
  xlab("Individual") +
  #
  guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) +
  ylab("Relative Abundance (Class > 2%) \n") +
  ggtitle("Class Composition of Nematostella vectensis \n Bacterial Communities by Tissue Compartment")
```
```{r, echo=FALSE, include=FALSE, warning=FALSE, message = FALSE}
Class_comp <- ggplot(Nvec_class, aes(x = Individual, y = Abundance, fill = Class)) + 
  facet_grid(Tissue~., labeller = labeller(Tissue = Tissue.labs)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = class_colors) +
  scale_x_discrete(
    breaks = c("1", "2", "3", "4", "5", "6"),
    labels = c("1", "2", "3", "4", "5", "6"), 
    drop = FALSE
  ) +
  # Remove x axis title
  xlab("Individual") +
  #
  guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) +
  ylab("Relative Abundance (Class > 2%) \n") +
  ggtitle("Class Composition of Nematostella vectensis \n Bacterial Communities by Tissue Compartment")

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/Revised_Class_comp.tiff", units="in", width=5, height=5, res=300)
plot(Class_comp)
dev.off()
```
## Order
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Nvec_order <- ps %>%
  tax_glom(taxrank = "Order") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.02) %>%                         # Filter out low abundance taxa
  arrange(Order) 
Order_colors <- c(
  "cyan3", "mediumorchid4", "dodgerblue", "mediumpurple1", "mediumorchid3", "tan1", "slateblue3", "tomato", "slateblue1","aquamarine", "mediumorchid1", "tan4", "darkcyan", "dodgerblue2","cadetblue3", "mediumpurple2","mediumpurple3", "mediumpurple4","palegreen3", "darkslategray4"
)
ggplot(Nvec_order, aes(x = Individual, y = Abundance, fill = Order)) + 
  facet_grid(Tissue~., labeller = labeller(Tissue = Tissue.labs)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = Order_colors) +
  scale_x_discrete(
    breaks = c("1", "2", "3", "4", "5", "6"),
    labels = c("1", "2", "3", "4", "5", "6"), 
    drop = FALSE
  ) +
  # Remove x axis title
  xlab("Individual") +
  #
  guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) +
  ylab("Relative Abundance (Order > 2%) \n") +
  ggtitle("Order Composition of Nematostella vectensis \n Bacterial Communities by Tissue Compartment")

```
```{r, echo=FALSE, include=FALSE, warning=FALSE, message = FALSE}
Order_comp <- ggplot(Nvec_order, aes(x = Individual, y = Abundance, fill = Order)) + 
  facet_grid(Tissue~., labeller = labeller(Tissue = Tissue.labs)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = Order_colors) +
  scale_x_discrete(
    breaks = c("1", "2", "3", "4", "5", "6"),
    labels = c("1", "2", "3", "4", "5", "6"), 
    drop = FALSE
  ) +
  # Remove x axis title
  xlab("Individual") +
  #
  guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) +
  ylab("Relative Abundance (Order > 2%) \n") +
  ggtitle("Order Composition of Nematostella vectensis \n Bacterial Communities by Tissue Compartment")

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/Order_comp_new.tiff", units="in", width=7, height=5, res=300)
plot(Order_comp)
dev.off()
```
## Family
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Nvec_Family <- ps %>%
  tax_glom(taxrank = "Family") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.05) %>%                         # Filter out low abundance taxa
  arrange(Family) 
Family_colors <- c(
  "#CBD588", "#5F7FC7", "orange","#DA5724", "#508578", "#CD9BCD",
   "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861", "chocolate1", "beige", "coral", "cyan1", "purple"
)
ggplot(Nvec_Family, aes(x = Individual, y = Abundance, fill = Family)) + 
  facet_grid(Tissue~., labeller = labeller(Tissue = Tissue.labs)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = Family_colors) +
  scale_x_discrete(
    breaks = c("1", "2", "3", "4", "5", "6"),
    labels = c("1", "2", "3", "4", "5", "6"), 
    drop = FALSE
  ) +
  # Remove x axis title
  xlab("Individual") +
  #
  guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) +
  ylab("Relative Abundance (Family > 5%) \n") +
  ggtitle("Family Composition of Nematostella vectensis \n Bacterial Communities by Tissue Compartment")

```
```{r, echo=FALSE, include=FALSE, warning=FALSE}
Family_comp <- ggplot(Nvec_Family, aes(x = Individual, y = Abundance, fill = Family)) + 
  facet_grid(Tissue~., labeller = labeller(Tissue = Tissue.labs)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = Family_colors) +
  scale_x_discrete(
    breaks = c("1", "2", "3", "4", "5", "6"),
    labels = c("1", "2", "3", "4", "5", "6"), 
    drop = FALSE
  ) +
  # Remove x axis title
  xlab("Individual") +
  #
  guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) +
  ylab("Relative Abundance (Family > 5%) \n") +
  ggtitle("Family Composition of Nematostella vectensis \n Bacterial Communities by Tissue Compartment")

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/Revised_Family_comp.tiff", units="in", width=7, height=5, res=300)
plot(Family_comp)
dev.off()
```
# Fixed PCoA
```{r, echo=FALSE, warning=FALSE, message = FALSE}
sample_data(ps)$Individual <- factor(
  sample_data(ps)$Individual, 
  levels = c("1", "2", "3", "4", "5", "6")
)
Nvec_pcoa <- ordinate(
  physeq = ps, 
  method = "PCoA", 
  distance = "bray"
)

plot_ordination(
  physeq = ps,
  ordination = Nvec_pcoa,
  color = "Individual",
  shape = "Tissue",
  title = "PCoA of Nematostella vectensis Bacterial Communities"
) + 
  scale_color_manual(values = c("#a65628", "red", "#ffae19",
    "#4daf4a", "#1919ff", "darkorchid3", "magenta")
  ) +
  geom_point(aes(color = Individual), alpha = 0.7, size = 4) +
  geom_point(colour = "grey90", size = 1.5) 
```
```{r, echo=FALSE, warning=FALSE, message = FALSE}
BrayPCoA_NVec <- plot_ordination(
  physeq = ps,
  ordination = Nvec_pcoa,
  color = "Individual",
  shape = "Tissue",
  title = "PCoA of Nematostella vectensis Bacterial Communities"
) + 
  scale_color_manual(values = c("#a65628", "red", "#ffae19",
    "#4daf4a", "#1919ff", "darkorchid3", "magenta")
  ) +
  geom_point(aes(color = Individual), alpha = 0.7, size = 4) +
  geom_point(colour = "grey90", size = 1.5) 
svg(filename="/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/BrayPCoA_NVec.svg")
plot(BrayPCoA_NVec)
dev.off()
png(filename="/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/BrayPCoA_NVec.png")
plot(BrayPCoA_NVec)
dev.off()
```

## PERMANOVA
```{r, echo=FALSE, warning=FALSE}
Nvec_bray <- phyloseq::distance(ps, method = "bray")

# make a data frame from the sample_data
sampledf <- data.frame(sample_data(ps))

# Adonis test
adonis(Nvec_bray ~ Tissue, data = sampledf)

```
## Pie Charts
```{r, echo=FALSE, include=FALSE, warning=FALSE, message = FALSE}
phyla_counts_tab <- otu_table(tax_glom(ps, taxrank="Phylum")) 
phyla_tax_vec <- as.vector(tax_table(tax_glom(ps, taxrank="Phylum"))[,2]) 
rownames(phyla_counts_tab) <- as.vector(phyla_tax_vec)
unclassified_tax_counts <- colSums(SV) - colSums(phyla_counts_tab)
phyla_and_unidentified_counts_tab <- rbind(phyla_counts_tab, "Unclassified"=unclassified_tax_counts)
temp_major_taxa_counts_tab <- phyla_and_unidentified_counts_tab[!row.names(phyla_and_unidentified_counts_tab) %in% "Proteobacteria", ]
class_counts_tab <- otu_table(tax_glom(ps, taxrank="Class"))
class_tax_phy_tab <- tax_table(tax_glom(ps, taxrank="Class"))
phy_tmp_vec <- class_tax_phy_tab[,2]
class_tmp_vec <- class_tax_phy_tab[,3]
rows_tmp <- row.names(class_tax_phy_tab)
class_tax_tab <- data.frame("Phylum"=phy_tmp_vec, "Class"=class_tmp_vec, row.names = rows_tmp)
proteo_classes_vec <- as.vector(class_tax_tab[class_tax_tab$Phylum == "Proteobacteria", "Class"])
rownames(class_counts_tab) <- as.vector(class_tax_tab$Class) 
proteo_class_counts_tab <- class_counts_tab[row.names(class_counts_tab) %in% proteo_classes_vec, ]
proteo_no_class_annotated_counts <- phyla_and_unidentified_counts_tab[row.names(phyla_and_unidentified_counts_tab) %in% "Proteobacteria", ] - colSums(proteo_class_counts_tab)
major_taxa_counts_tab <- rbind(temp_major_taxa_counts_tab, proteo_class_counts_tab, "Unresolved_Proteobacteria"=proteo_no_class_annotated_counts)
identical(colSums(major_taxa_counts_tab), colSums(SV))
major_taxa_proportions_tab <- apply(major_taxa_counts_tab, 2, function(x) x/sum(x)*100)
dim(major_taxa_proportions_tab)
temp_filt_major_taxa_proportions_tab <- data.frame(major_taxa_proportions_tab[apply(major_taxa_proportions_tab, 1, max) > 5, ])
dim(temp_filt_major_taxa_proportions_tab) 
filtered_proportions <- colSums(major_taxa_proportions_tab) - colSums(temp_filt_major_taxa_proportions_tab)
filt_major_taxa_proportions_tab <- rbind(temp_filt_major_taxa_proportions_tab, "Other"=filtered_proportions)
filt_major_taxa_proportions_tab_for_plot <- filt_major_taxa_proportions_tab
filt_major_taxa_proportions_tab_for_plot$Major_Taxa <- row.names(filt_major_taxa_proportions_tab_for_plot)
filt_major_taxa_proportions_tab_for_plot.g <- gather(filt_major_taxa_proportions_tab_for_plot, Sample, Proportion, -Major_Taxa)
head(filt_major_taxa_proportions_tab_for_plot.g)
head(filt_major_taxa_proportions_tab_for_plot)
sample_info_for_merge<-data.frame("Sample"=map$ID, "Tissue"=map$Tissue, "Individual"=map$Individual, stringsAsFactors=F)
filt_major_taxa_proportions_tab_for_plot.g2 <- merge(filt_major_taxa_proportions_tab_for_plot.g, sample_info_for_merge)
```

## Capitulum
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Cap_sample_IDs <- row.names(map)[map$Tissue == "capitulum"]
filt_major_taxa_proportions_cap_only_tab_for_plot.g <- filt_major_taxa_proportions_tab_for_plot.g2[filt_major_taxa_proportions_tab_for_plot.g2$Sample %in% Cap_sample_IDs, ]
Cap_sample_major_taxa_proportion_tab <- acast(filt_major_taxa_proportions_cap_only_tab_for_plot.g[, c(1:3)], Major_Taxa ~ Sample)
Cap_sample_summed_major_taxa_proportions_vec <- rowSums(Cap_sample_major_taxa_proportion_tab)
Cap_sample_major_taxa_summary_tab <- data.frame("Major_Taxa"=names(Cap_sample_summed_major_taxa_proportions_vec), "Proportion"=Cap_sample_summed_major_taxa_proportions_vec, row.names=NULL)
ggplot(data.frame(Cap_sample_major_taxa_summary_tab), aes(x="Capitulum samples", y=Proportion, fill=Major_Taxa)) + 
  geom_bar(width=1, stat="identity") +
  coord_polar("y") +
  scale_fill_viridis(discrete=TRUE) +
  ggtitle("Capitulum") +
  theme_void() +
  theme(plot.title = element_text(hjust=0.5), legend.title=element_blank())
```
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Cap_pie <- ggplot(data.frame(Cap_sample_major_taxa_summary_tab), aes(x="Capitulum samples", y=Proportion, fill=Major_Taxa)) + 
  geom_bar(width=1, stat="identity") +
  coord_polar("y") +
  scale_fill_viridis(discrete=TRUE) +
  ggtitle("Capitulum") +
  theme_void() +
  theme(plot.title = element_text(hjust=0.5), legend.title=element_blank())

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/Cap_pie.tiff", units="in", width=6, height=5, res=300)
plot(Cap_pie)
dev.off()
```

## Mesenteries
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Mes_sample_IDs <- row.names(map)[map$Tissue == "mesenteries"]
filt_major_taxa_proportions_mes_only_tab_for_plot.g <- filt_major_taxa_proportions_tab_for_plot.g2[filt_major_taxa_proportions_tab_for_plot.g2$Sample %in% Mes_sample_IDs, ]
Mes_sample_major_taxa_proportion_tab <- acast(filt_major_taxa_proportions_mes_only_tab_for_plot.g[, c(1:3)], Major_Taxa ~ Sample)
Mes_sample_summed_major_taxa_proportions_vec <- rowSums(Mes_sample_major_taxa_proportion_tab)
Mes_sample_major_taxa_summary_tab <- data.frame("Major_Taxa"=names(Mes_sample_summed_major_taxa_proportions_vec), "Proportion"=Mes_sample_summed_major_taxa_proportions_vec, row.names=NULL)
ggplot(data.frame(Mes_sample_major_taxa_summary_tab), aes(x="Mesentery samples", y=Proportion, fill=Major_Taxa)) + 
  geom_bar(width=1, stat="identity") +
  coord_polar("y") +
  scale_fill_viridis(discrete=TRUE) +
  ggtitle("Mesenteries") +
  theme_void() +
  theme(plot.title = element_text(hjust=0.5), legend.title=element_blank())
```
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Mes_pie <- ggplot(data.frame(Mes_sample_major_taxa_summary_tab), aes(x="Mesentery samples", y=Proportion, fill=Major_Taxa)) + 
  geom_bar(width=1, stat="identity") +
  coord_polar("y") +
  scale_fill_viridis(discrete=TRUE) +
  ggtitle("Mesenteries") +
  theme_void() +
  theme(plot.title = element_text(hjust=0.5), legend.title=element_blank())

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/Mes_pie.tiff", units="in", width=6, height=5, res=300)
plot(Mes_pie)
dev.off()
```

## Physa
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Phy_sample_IDs <- row.names(map)[map$Tissue == "physa"]
filt_major_taxa_proportions_phy_only_tab_for_plot.g <- filt_major_taxa_proportions_tab_for_plot.g2[filt_major_taxa_proportions_tab_for_plot.g2$Sample %in% Phy_sample_IDs, ]
Phy_sample_major_taxa_proportion_tab <- acast(filt_major_taxa_proportions_phy_only_tab_for_plot.g[, c(1:3)], Major_Taxa ~ Sample)
Phy_sample_summed_major_taxa_proportions_vec <- rowSums(Phy_sample_major_taxa_proportion_tab)
Phy_sample_major_taxa_summary_tab <- data.frame("Major_Taxa"=names(Phy_sample_summed_major_taxa_proportions_vec), "Proportion"=Phy_sample_summed_major_taxa_proportions_vec, row.names=NULL)
ggplot(data.frame(Phy_sample_major_taxa_summary_tab), aes(x="Physa samples", y=Proportion, fill=Major_Taxa)) + 
  geom_bar(width=1, stat="identity") +
  coord_polar("y") +
  scale_fill_viridis(discrete=TRUE) +
  ggtitle("Physa") +
  theme_void() +
  theme(plot.title = element_text(hjust=0.5), legend.title=element_blank())
```
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Phy_pie <- ggplot(data.frame(Phy_sample_major_taxa_summary_tab), aes(x="Physa samples", y=Proportion, fill=Major_Taxa)) + 
  geom_bar(width=1, stat="identity") +
  coord_polar("y") +
  scale_fill_viridis(discrete=TRUE) +
  ggtitle("Physa") +
  theme_void() +
  theme(plot.title = element_text(hjust=0.5), legend.title=element_blank())

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/Phy_pie.tiff", units="in", width=6, height=5, res=300)
plot(Phy_pie)
dev.off()
```
# Betadisperser
```{r, echo=FALSE, warning=FALSE}
DESeq_map <-read.table("/Users/anthonybonacolta/Box/Nvec_Microbiome/R_data/DESeq_map.txt", sep ="\t", row.names = 1, header = TRUE)
deseq_counts <- DESeqDataSetFromMatrix(SV[,-19], colData =DESeq_map, design = ~Tissue)
deseq_counts_vst <- varianceStabilizingTransformation(deseq_counts)
vst_trans_count_tab <- assay(deseq_counts_vst)
euc_dist <- dist(t(vst_trans_count_tab))
euc_clust <- hclust(euc_dist, method="ward.D2")
plot(euc_clust)
```
```{r}
anova(betadisper(euc_dist, map$Tissue))
adonis(euc_dist~map$Tissue)
```
## Visualize PERMANOVA
```{r, echo=FALSE}
vst_pcoa <- ordinate(ps, method="MDS", distance="euclidean")

plot_ordination(
  physeq = ps,
  ordination = vst_pcoa,
  color = "Individual",
  shape = "Tissue",
  title = "PCoA of Nematostella vectensis Bacterial Communities"
) + 
  scale_color_manual(values = c("#a65628", "red", "#ffae19",
    "#4daf4a", "#1919ff", "darkorchid3", "magenta")
  ) +
  geom_point(aes(color = Individual), alpha = 0.7, size = 4) +
  geom_point(colour = "grey90", size = 1.5) + 
  annotate("text", x=40000, y=40000, label="Tissue Regions") +
  annotate("text", x=40000, y=36000, label="Permutational ANOVA = 0.041*")

```
```{r, echo=FALSE}
PCoA_Euc <- plot_ordination(
  physeq = ps,
  ordination = vst_pcoa,
  color = "Individual",
  shape = "Tissue",
  title = "PCoA of Nematostella vectensis Bacterial Communities"
) + 
  scale_color_manual(values = c("#a65628", "red", "#ffae19",
    "#4daf4a", "#1919ff", "darkorchid3", "magenta")
  ) +
  geom_point(aes(color = Individual), alpha = 0.7, size = 4) +
  geom_point(colour = "grey90", size = 1.5) + 
  annotate("text", x=40000, y=40000, label="Tissue Regions") +
  annotate("text", x=40000, y=36000, label="Permutational ANOVA = 0.041*")

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/PCoA_Euc.tiff", units="in", width=7, height=5, res=300)
plot(PCoA_Euc)
dev.off()
```

# MicroBiomeSeq Analysis for Alpha and Beta Diversity Measures

```{r, echo=FALSE, warning=FALSE, message = FALSE, include=FALSE}
library(devtools)
library(phyloseq)
library(microbiomeSeq)
library(ggplot2)
physeq = ps
```

# Alpha Diversity with ANOVA (P<0.05)
```{r, echo=FALSE, warning=FALSE, message = FALSE}
Alpha_ANOVA <- plot_anova_diversity(physeq, method = c("richness", "simpson", "shannon"), 
    grouping_column = "Tissue", pValueCutoff = 0.05)
print(Alpha_ANOVA)

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/Alpha_ANOVA.tiff", units="in", width=7, height=5, res=300)
plot(Alpha_ANOVA)
dev.off()
```

## Differentially abundant taxa 
### Between Tissue compartmen, corresponding adjusted p-values and rank of importance as detected by random forest classifier.

```{r, echo=FALSE, warning=FALSE, message = FALSE, fig.width=5,fig.height=5}
physeq2 <- taxa_level(physeq, "Phylum")
deseq_sig <- differential_abundance(physeq2, grouping_column = "Tissue", output_norm = "log-relative", 
    pvalue.threshold = 0.05, lfc.threshold = 0, filename = F)
diff_Phy <- plot_signif(deseq_sig$plotdata, top.taxa = 5)
print(diff_Phy)

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/diff_Phy.tiff", units="in", width=7, height=8, res=300)
plot(diff_Phy)
dev.off()
```

#### Very little or no Lentisphaerae in Capitulum Samples

## Mean Decrease Accuracy
```{r, echo=FALSE, warning=FALSE, message = FALSE}
MDA <- plot_MDA(deseq_sig$importance)
print(MDA)

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/MDA.tiff", units="in", width=7, height=5, res=300)
plot(MDA)
dev.off()
```

### Factor in Abundance 
```{r, echo=FALSE, warning=FALSE, message = FALSE}
MA <- plot_MA(deseq_sig$SignFeaturesTable)
print(MA$lfcplot)

tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/MA.tiff", units="in", width=7, height=8, res=300)
plot(MA$lfcplot)
dev.off()
```
 
 
 **Down reg are more common in Cap, Up reg are more common in other compartments**
 
 
```{r}
library("ampvis2")
if(!require("devtools"))
  install.packages("devtools")
#source the phyloseq_to_ampvis2() function from the gist
devtools::source_gist("8d0ca4206a66be7ff6d76fc4ab8e66c6")
ampvis2_obj <- phyloseq_to_ampvis2(ps)
```
 
 
```{r}
hm <- amp_heatmap(ampvis2_obj, group_by = "Tissue", tax_aggregate = "Family", tax_show = 9)+theme_classic()+labs(title="Read Abundance for Top Bacterial Families", subtitle="Divided by Tissue Compartment")
tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/hm.tiff", units="in", width=8, height=5, res=300)
plot(hm)
dev.off()
```
 
```{r}
library(ALDEx2);
packageVersion("ALDEx2")
library(CoDaSeq)
packageVersion("CoDaSeq")

d.czm.all <- cmultRepl(SV[1:18], method="CZM", label=0)
d.clr.all <- codaSeq.clr(d.czm.all)
E.clr <- t(d.clr.all)
d.pcx <- prcomp(E.clr)

dist.clr <- dist(E.clr)
```
```{r}
tiss <- map$Tissue
ano <- anosim(dist.clr, tiss, permutations=9999)
ano
```
```{r}
plot(ano)
```
```{r}
df_out <- as.data.frame(d.pcx$x)
p<-ggplot(df_out,aes(x=PC1,y=PC2,fill=factor(map$Individual),shape=map$Tissue))
p<-p+geom_point(size=3, position=position_jitter(h=0.1, w=0.1))+
  scale_shape_manual(values=c(21,24,22), name = "Tissue") +scale_fill_viridis_d(option="magma", name = "Individual") +
  guides(fill = guide_legend(override.aes=list(shape=21))) + ggtitle("PCA of Nematostella vectensis Bacterial Communities") + annotate("text", x=35, y=75, label="Tissue ANOSIM = 0.1483 \n (9999 Permutations)") +theme_bw()
p

#tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/PCA_Ait.tiff", units="in", width=7, height=5, res=300)
#plot(p)
#dev.off()
```

# Differential Abundance

```{r}
library("microbiome"); packageVersion("microbiome")
library(exactRankTests)
library(nlme)
library(dplyr)
library(ggplot2)
library(compositions)
library(ancom.R)
```
```{r}
ps_filtered <- prune_taxa(taxa_sums(ps) > 6, ps)
ps_filtered
ps_clr <- microbiome::transform(ps_filtered, 'clr', shift = 1)
```

```{r}
feature_table_pre_process = function(feature_table, meta_data, sample_var, group_var = NULL, 
                                     out_cut = 0.05, zero_cut = 0.90, lib_cut = 200, neg_lb){
  feature_table = data.frame(feature_table, check.names = FALSE)
  meta_data = data.frame(meta_data, check.names = FALSE)
  # Drop unused levels
  meta_data[] = lapply(meta_data, function(x) if(is.factor(x)) factor(x) else x)
  # Match sample IDs between metadata and feature table
  sample_ID = intersect(meta_data[, sample_var], colnames(feature_table))
  feature_table = feature_table[, sample_ID]
  meta_data = meta_data[match(sample_ID, meta_data[, sample_var]), ]
  
  # 1. Identify outliers within each taxon
  if (!is.null(group_var)) {
    group = meta_data[, group_var]
    z = feature_table + 1 # Add pseudo-count (1) 
    f = log(z); f[f == 0] = NA; f = colMeans(f, na.rm = T)
    f_fit = lm(f ~ group)
    e = rep(0, length(f)); e[!is.na(group)] = residuals(f_fit)
    y = t(t(z) - e)
    
    outlier_check = function(x){
      # Fitting the mixture model using the algorithm of Peddada, S. Das, and JT Gene Hwang (2002)
      mu1 = quantile(x, 0.25, na.rm = T); mu2 = quantile(x, 0.75, na.rm = T)
      sigma1 = quantile(x, 0.75, na.rm = T) - quantile(x, 0.25, na.rm = T); sigma2 = sigma1
      pi = 0.75
      n = length(x)
      epsilon = 100
      tol = 1e-5
      score = pi*dnorm(x, mean = mu1, sd = sigma1)/((1 - pi)*dnorm(x, mean = mu2, sd = sigma2))
      while (epsilon > tol) {
        grp1_ind = (score >= 1)
        mu1_new = mean(x[grp1_ind]); mu2_new = mean(x[!grp1_ind])
        sigma1_new = sd(x[grp1_ind]); if(is.na(sigma1_new)) sigma1_new = 0
        sigma2_new = sd(x[!grp1_ind]); if(is.na(sigma2_new)) sigma2_new = 0
        pi_new = sum(grp1_ind)/n
        
        para = c(mu1_new, mu2_new, sigma1_new, sigma2_new, pi_new)
        if(any(is.na(para))) break
        
        score = pi_new * dnorm(x, mean = mu1_new, sd = sigma1_new)/
          ((1-pi_new) * dnorm(x, mean = mu2_new, sd = sigma2_new))
        
        epsilon = sqrt((mu1 - mu1_new)^2 + (mu2 - mu2_new)^2 + 
                         (sigma1 - sigma1_new)^2 + (sigma2 - sigma2_new)^2 + (pi - pi_new)^2)
        mu1 = mu1_new; mu2 = mu2_new; sigma1 = sigma1_new; sigma2 = sigma2_new; pi = pi_new
      }
      
      if(mu1 + 1.96 * sigma1 < mu2 - 1.96 * sigma2){
        if(pi < out_cut){
          out_ind = grp1_ind
        }else if(pi > 1 - out_cut){
          out_ind = (!grp1_ind)
        }else{
          out_ind = rep(FALSE, n)
        }
      }else{
        out_ind = rep(FALSE, n)
      }
      return(out_ind)
    }
    out_ind = matrix(FALSE, nrow = nrow(feature_table), ncol = ncol(feature_table))
    out_ind[, !is.na(group)] = t(apply(y, 1, function(i) 
      unlist(tapply(i, group, function(j) outlier_check(j)))))
    
    feature_table[out_ind] = NA
  }
  
  # 2. Discard taxa with zeros  >=  zero_cut
  zero_prop = apply(feature_table, 1, function(x) sum(x == 0, na.rm = T)/length(x[!is.na(x)]))
  taxa_del = which(zero_prop >= zero_cut)
  if(length(taxa_del) > 0){
    feature_table = feature_table[- taxa_del, ]
  }
  
  # 3. Discard samples with library size < lib_cut
  lib_size = colSums(feature_table, na.rm = T)
  if(any(lib_size < lib_cut)){
    subj_del = which(lib_size < lib_cut)
    feature_table = feature_table[, - subj_del]
    meta_data = meta_data[- subj_del, ]
  }
  
  # 4. Identify taxa with structure zeros
  if (!is.null(group_var)) {
    group = factor(meta_data[, group_var])
    present_table = as.matrix(feature_table)
    present_table[is.na(present_table)] = 0
    present_table[present_table != 0] = 1
    
    p_hat = t(apply(present_table, 1, function(x)
      unlist(tapply(x, group, function(y) mean(y, na.rm = T)))))
    samp_size = t(apply(feature_table, 1, function(x)
      unlist(tapply(x, group, function(y) length(y[!is.na(y)])))))
    p_hat_lo = p_hat - 1.96 * sqrt(p_hat * (1 - p_hat)/samp_size)
    
    struc_zero = (p_hat == 0) * 1
    # Whether we need to classify a taxon into structural zero by its negative lower bound?
    if(neg_lb) struc_zero[p_hat_lo <= 0] = 1
    
    # Entries considered to be structural zeros are set to be 0s
    struc_ind = struc_zero[, group]
    feature_table = feature_table * (1 - struc_ind)
    
    colnames(struc_zero) = paste0("structural_zero (", colnames(struc_zero), ")")
  }else{
    struc_zero = NULL
  }
  
  # 5. Return results
  res = list(feature_table = feature_table, meta_data = meta_data, structure_zeros = struc_zero)
  return(res)
}
```

```{r}
feature_table = ps@otu_table; sample_var = "Sample"; group_var = "Tissue"
meta_data = ps@sam_data
out_cut = 0.05; zero_cut = 0.90; lib_cut = 200; neg_lb = FALSE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var, 
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
```

```{r}
main_var = "Tissue"; p_adj_method = "BH"; alpha = 0.05
#feature_table = t(SV)
#write.csv(feature_table,"/Users/anthonybonacolta/Box/Nvec_Microbiome/feature_table.csv")
feature_table <- read.table("/Users/anthonybonacolta/Box/Nvec_Microbiome/feature_table.csv",sep = ",", row.names = 1, header = T)
res = ANCOM(feature_table, sig=0.05,multcorr = 3, tau = 0.02, theta = 0.1)
plot_ancom(res)
write.csv(res$detected, "/Users/anthonybonacolta/Box/Nvec_Microbiome/ANCOM.csv")
```

## New PCOA plots

# Filter low abundance OTUs
```{r}
ps_filtered <- prune_taxa(taxa_sums(ps) > 6, ps)
ps_filtered
```

# CLR transform
```{r}
ps_clr <- microbiome::transform(ps_filtered, 'clr', shift = 1)
```
# Ordinate
```{r}
sample_data(ps_clr)$Individual <- factor(sample_data(ps_clr)$Individual, levels = c("1", "2", "3", "4", "5", "6"))
                                         
psr_clr.ord <- ordinate(ps_clr, "RDA", "euclidean")
PCA = plot_ordination(ps_clr, psr_clr.ord, 
                                shape="Individual",
                                color="Tissue",
                                title="Aitchison Distance PCA") +
  scale_color_manual(values = c("darkgreen", "navyblue", "purple")) + scale_shape_manual(values = c(0,1,2,5,6,7)) + theme_bw() +
  geom_point(aes(color = Tissue), alpha = 0.7, size = 5)

PCA
```

```{r}
dist.clr <- dist(otu_table(ps_clr)[0:18], method ="euclidean")
anosim(dist.clr, ps_clr@sam_data$Tissue, permutations = 999)
```
# Betadisperser
```{r, warning=FALSE}
DESeq_map <-read.table("/Users/anthonybonacolta/Box/Nvec_Microbiome/R_data/DESeq_map.txt", sep ="\t", row.names = 1, header = TRUE)
deseq_counts <- DESeqDataSetFromMatrix(SV[,-19], colData =DESeq_map, design = ~Tissue)
deseq_counts_vst <- varianceStabilizingTransformation(deseq_counts)
vst_trans_count_tab <- assay(deseq_counts_vst)
euc_dist <- dist(t(vst_trans_count_tab))
euc_clust <- hclust(euc_dist, method="ward.D2")
```


## Visualize PCoA
```{r}
sample_data(ps)$Individual <- factor(
  sample_data(ps)$Individual, 
  levels = c("1", "2", "3", "4", "5", "6")
)
Nvec_pcoa <- ordinate(
  physeq = ps, 
  method = "PCoA", 
  distance = "bray"
)

plot_ordination(
  physeq = ps,
  ordination = Nvec_pcoa,
  color = "Tissue",
  label = "Individual",
  title = "Bray-Curtis PCoA"
) + 
  scale_color_manual(values = c("darkgreen", "navyblue", "purple")
  ) +  theme_bw() +
  geom_point(aes(color = Tissue), alpha = 0.7, size = 5, label= "Individual")
```

```{r}
tiff("/Users/anthonybonacolta/Box/Nvec_Microbiome/Figures/TIFF/Bray_PCoA_new.tiff", units="in", width=7, height=5, res=300)
plot_ordination(
  physeq = ps,
  ordination = Nvec_pcoa,
  color = "Tissue",
  label = "Individual",
  title = "Bray-Curtis PCoA"
) + 
  scale_color_manual(values = c("darkgreen", "navyblue", "purple")
  ) +  theme_bw() +
  geom_point(aes(color = Tissue), alpha = 0.7, size = 5, label= "Individual")
dev.off()
```









